Draft: Notes on the revised Hidden weight function

### 2024-09-25 RevisedHiddenWeight.tex
ok

%\begin{definition}[Algebraic Immunity] \label{def:ai}
%	The algebraic immunity of a Boolean function $f\in \Bfn{n}$, denoted as $\AI(f)$, is defined as:
%	\[ \AI(f) = \min_{g \neq 0}\{ \degg(g) \; | \; fg = 0 \; \text{or} \; (f + 1)g = 0 \}{,} \]
%	where $\degg(g)$ is the algebraic degree of $g$.
%	The function $g$ is called an annihilator of $f$ (or $f + 1$). 
%	Additionally we denote $\AN(f) = \min_{g \neq 0}\{ \degg(g) \; | \; fg = 0\}$.
%\end{definition}

@@

The $n$-variable \emph{Boolean symmetric functions} are those that are constant on each slice $\Ekn{k}{n}$ for $k\in [0,n]$. 
This class has been assiduously studied in the context of cryptography, see \eg \cite{IEEE:Carlet04,IEEE:CanVid05,INDO:BraPre05,DM:SarMai07,IEEE:QFLW09,CCDS:Meaux21,IEEE:CarMea21}.
In this paper we mainly consider two families of symmetric functions, which are both bases of the symmetric functions:

@@

@@

Accordingly, $f$ is balanced when $n=0\mod 4$ and unbalanced when $n=2 \mod 4$.

@@

@@

\begin{proposition}[Algebraic immunity of $f$]
	Let $n\in \N$ be even, $f$ defined as
	$f(x_1,\cdots,x_n)=\left(\sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}\right) + \sum_{k=1}^n \phikn{k}{n} x_k$ and $h$ denote the $n$-variable HWBF, the following holds:
	\[\AI(f)\ge \AI(h)-2.\]
\end{proposition}
\begin{proof}
	content...
\end{proof}

>>>>

%\begin{definition}[Algebraic Immunity] \label{def:ai}
%	The algebraic immunity of a Boolean function $f\in \Bfn{n}$, denoted as $\AI(f)$, is defined as:
%	\[ \AI(f) = \min_{g \neq 0}\{ \degg(g) \; | \; fg = 0 \; \text{or} \; (f + 1)g = 0 \}{,} \]
%	where $\degg(g)$ is the algebraic degree of $g$.
%	The function $g$ is called an annihilator of $f$ (or $f + 1$). 
%	Additionally we denote $\AN(f) = \min_{g \neq 0}\{ \degg(g) \; | \; fg = 0\}$.
%\end{definition}
\begin{definition}[Nonlinearity and weightwise nonlinearity] \label{def:nl}
	The nonlinearity $\NL(f)$ of a Boolean function $f\in \BN$, where $n$ is a positive integer, is the minimum Hamming distance between $f$ and all the affine functions in $\BN$:
	\[ \NL(f) = \min_{g,\, \degg(g)\leq 1} \{ \hd(f,g) \}{,} \]
	where $g(x)=a\cdot x+\varepsilon$, $a\in \F_2^n, \varepsilon\in \F_2$. 	
	The nonlinearity can also be defined from the Walsh transform:
	\[ \NL(f) = 2^{n-1}- \frac{1}{2} \max_{a\in \F_2^n}{|\wt{f}(a)|}. \]
\end{definition}

\begin{definition}[Algebraic immunity] \label{def:ai}
	The algebraic immunity of a Boolean function $f\in \Bfn{n}$, denoted as $\AI(f)$, is defined as:
	\[ \AI(f) = \min_{g \neq 0}\{ \degg(g) \; | \; fg = 0 \; \text{or} \; (f + 1)g = 0 \}{,} \]
	where $\degg(g)$ is the algebraic degree of $g$.
	The function $g$ is called an annihilator of $f$ (or $f + 1$). 
\end{definition}

@@

The $n$-variable \emph{Boolean symmetric functions} are those that are constant on each slice $\Ekn{k}{n}$ for $k\in [0,n]$. 
This class has been assiduously studied in the context of cryptography, see \eg \cite{IEEE:Carlet04,IEEE:CanVid05,INDO:BraPre05,DM:SarMai07,IEEE:QFLW09,Latin:Meaux19,CCDS:Meaux21,IEEE:CarMea21}.
In this paper we mainly consider two families of symmetric functions, which are both bases of the symmetric functions:

@@

\pmnote{Introduce formally HWBF, and weightwise degree d functions}

@@

Accordingly, $f$ is balanced when $n \equiv 0\mod 4$ and unbalanced when $n\equiv 2 \mod 4$.

@@

\pmnote{@Tim, could you do a (summary) graph with the bound it gives for the nonlinearity? (Definition from the Walsh transform in the preliminaries)

Could you add curves for the majority function, $n$ even $\NL=2^{n-1}- 1/2 \binom{n}{n/2}$ , and HWBF  $n$ even $\NL=2^{n-1}- 2 \binom{n-2}{(n-2)/2}$ \cite{DAM:WCST14} Theorem 3.

}

@@

\begin{proposition}[Algebraic immunity of $f$]
	Let $n\in \N$ be even, $f$ defined as
	$f(x_1,\cdots,x_n)=\left(\sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}\right) + \sum_{k=1}^n \phikn{k}{n} x_k$ and $h$ denote the $n$-variable HWBF, the following holds:
	\[\AI(f)\ge \AI(h)-2.\]
\end{proposition}
\begin{proof}
	Let denote by $q$ the quadratic function $\sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}$, it allows to write $h=f+q$. 
	Let $g\neq 0$ be an annihilator of $f+ \varepsilon$ of minimum degree (where $\varepsilon \in \{0,1\}$), then the following holds:
	\[g (1+q) (h+\varepsilon) = g(1+q) (f+\varepsilon + q)=0, \]
	hence $g (1+q)$ is an annihilator of $h+ \varepsilon$. 
	If $g (1+q)\ne 0$ then $\degg(g(1+q)) \le \degg(g)+2$ and we obtain $\degg(g)\ge \AI(h) -2$. 
	Otherwise, $g (1+q)= 0$ implies that $g$ annihilates $1+q$ and therefore $0=g (f+\varepsilon + 1 +q)=g(h+ \varepsilon +1)$ that is $\degg(g)\ge \AI(h)$. 
	Summarizing, for any non null annihilator $g$ of $f$ or $f+1$ we get $\degg(g)\ge \AI(h)-2$, it allows to conclude $\AI(f)\ge \AI(h)-2$.

From~\cite{DAM:WCST14} Theorem 4, the algebraic immunity of the HWBF is at least $\lfloor n/3\rfloor +1$, which leads to $\lfloor n/3\rfloor -1$ for $f$.
